---
title: "Home Credit Default Risk Modeling"
author: "Joel Jorgensen"
date: today
format: 
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    embed-resources: true
execute:
  include: true
  eval: true    
  warning: false
  message: false

---
# Introduction
Home Credit is a multi national financial services provider offering a variety of loan products, they aim to make borrowing more accessible by offering loans to those customers who are typically under served,the under or unbanked.

A big challenge that arises from this is that  some customers lack a traditional credit score, this makes it more difficult to determine credit worthiness and decide who to extend loans to.

The goal of this project is to build a predictive model on which customers will default or face repayment problems, and to maximize the accuracy by using the data available at the time of their application and supplementing when possible from a few additional available data sources



# Load Data and Setup

```{r library}
pacman::p_load(tidyverse, caret, janitor, skimr, recipes, themis, rlang, tidymodels, here,caretEnsemble)
```

Data is all loaded in. This is 5 total data sets. The main set is the application data, then there are four additional data sets to supplement and try and improve the model accuracy. 
```{r load data application train}

dir<- getwd()
setwd(dir)

#load file - assumes you have copy of file in working directory
hc_train <-read.csv("application_train.csv", stringsAsFactors = TRUE)
application_train <- read.csv("application_train.csv")

#standardize variable names
hc_train <- hc_train |>
  rename_with(tolower)

#convert target and select numeric indicators to factors
hc_train <- hc_train |>
  mutate(
    target = as.factor(target),
    across(starts_with("flag"),as.factor),
    across(starts_with("reg_"),as.factor),
    across(starts_with("live_"),as.factor)
    )
```

```{r load data bureau}
#load file - assumes you have copy of file in working directory
bureau_data <- read.csv("bureau.csv", stringsAsFactors = TRUE)

#standardize variable names
bureau_data <- bureau_data |>
  rename_with(tolower)

```

```{r load data installment}
#load file - assumes you have copy of file in working directory
installment_data <- read.csv("installments_payments.csv", stringsAsFactors = TRUE)

#standardize variable names
installment_data <- installment_data |>
  rename_with(tolower)

```

```{r load credit card data}
#load file - assumes you have copy of file in working directory
credit_card_data <- read.csv("credit_card_balance.csv", stringsAsFactors = TRUE)

#filter to only include selected columns
credit_card_data <- credit_card_data |>
  select(
    SK_ID_CURR,
    MONTHS_BALANCE,
    AMT_BALANCE,
    AMT_CREDIT_LIMIT_ACTUAL,
    AMT_INST_MIN_REGULARITY,
    AMT_PAYMENT_CURRENT,
    AMT_PAYMENT_TOTAL_CURRENT,
    SK_DPD
  )
```

```{r load pos bal data}
#load file - assumes you have copy of file in working directory
pos_cash_data <-read.csv("POS_CASH_balance.csv", stringsAsFactors = TRUE)

#filter to only include selected columns
pos_cash_data <- pos_cash_data |>
  select(SK_ID_CURR,
         MONTHS_BALANCE,
         CNT_INSTALMENT,
         CNT_INSTALMENT_FUTURE,
         SK_DPD
  )
```


#EDA

## Target Variable
First we look at the target variable we can see customers have repayment difficulties about 8% of the time, if we were to predict in all cases that customers wouldn't have a repayment issue we would be correct 92% of the time. This indicates a noticeable class in balance. 
```{r class balance}
round(mean(application_train$TARGET) * 100,2)
```

## Further Feature Exploration

Next we will look at the categorical variables in comparison to the target variable. 
```{r}
application_train_clean <- application_train
#get target and char columns
application_train_clean |> select_if(is.character) -> char_df
char_df <- char_df |> mutate(TARGET = application_train_clean$TARGET)
```

For NAME_CONTRACT_TYPE we see a higher proportion of payment issues with Cash compared to Resolving loans.  
```{r name_contract_type}

char_df |> ggplot(aes(x = NAME_CONTRACT_TYPE, fill = factor(TARGET))) + 
  geom_bar(position = "fill") + labs(title = "Proportion of Target by Contract Type",
    y = "Proportion",
    x = "Contract Type",
    fill = "Target Variable")
  scale_y_continuous(labels = scales::percent)
```
For applicants who own cars and real estate they have a slightly lower proportion of repayment difficulty, the relative proportions look quite similar between the two features. 
```{r own car}

char_df |> ggplot(aes(x = FLAG_OWN_CAR, fill = factor(TARGET))) + 
  geom_bar(position = "fill") + labs(title = "Proportion of Target by Own CAR",
    y = "Proportion",
    x = "Contract Type",
    fill = "Target Variable")
  scale_y_continuous(labels = scales::percent)
```


If we look at NAME Income TYPE, we see that those on maternity leave and unemployed have much higher proportions of payment difficulties. For these two cases this would be a very strong predictor. 
```{r income type}
char_df |> ggplot(aes(x = NAME_INCOME_TYPE, fill = factor(TARGET))) + 
  geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 45)) + labs(title = "Proportion of Target Name Income Type",
    y = "Proportion",
    x = "Contract Type",
    fill = "Target Variable")
  scale_y_continuous(labels = scales::percent) 
```


We see a slightly higher of payment problems proportionally with those who rent apartments or live with their parents. 
```{r housing type}
char_df |> ggplot(aes(x = NAME_HOUSING_TYPE, fill = factor(TARGET))) + 
  geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 45)) + labs(title = "Proportion of Target Name Housing Type",
    y = "Proportion",
    x = "Contract Type",
    fill = "Target Variable")
  scale_y_continuous(labels = scales::percent) 
```

We do see some variation based on different occupation types, with occupations like drivers and low skilled laborers having the high proportion of payment difficulties. This could be explained largely by these occupations having the lowest pay. 
```{r occupation type}
char_df |> ggplot(aes(x = OCCUPATION_TYPE, fill = factor(TARGET))) + 
  geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 45)) + labs(title = "Proportion of Target Name Housing Type",
    y = "Proportion",
    x = "Contract Type",
    fill = "Target Variable")
  scale_y_continuous(labels = scales::percent) 
```
If we look at average pay by occupation, we can see the lowest paying jobs on average tend to have higher proportions of repayment difficulties in the chart above. This will be worth keeping in mind when building a model as these items are closely related.
```{r pay by occupation}
#average pay by occupation
application_train_clean |> group_by(OCCUPATION_TYPE) |> summarise(avg_inc = mean(AMT_INCOME_TOTAL)) |> arrange(avg_inc)
```


# Additional Data Source Preperation
Calculated columns are created for the additional data sources. Then for each an aggregation is performed to try and capture any relevant signal while getting it so their is one row for each client. 

## Bureau Data
```{r created aggregated bureau data}
#creates new data frame of aggregated bureau data
bureau_agg <- bureau_data |>
    group_by(sk_id_curr) |>
    summarise(
      bur_n_loans = n(),
      #status counts
      bur_n_active       = sum(credit_active == "Active", na.rm = TRUE),
      bur_n_closed       = sum(credit_active == "Closed", na.rm = TRUE),
      #min/max days since credit application
      bur_min_days_credit = suppressWarnings(min(days_credit, na.rm = TRUE)),
      bur_max_days_credit = suppressWarnings(max(days_credit, na.rm = TRUE)),
      #amount aggregates
      bur_sum_credit_sum     = sum(amt_credit_sum, na.rm = TRUE),
      bur_sum_credit_debt    = sum(amt_credit_sum_debt, na.rm = TRUE),
      bur_sum_credit_overdue = sum(amt_credit_sum_overdue, na.rm = TRUE),
      bur_sum_credit_limit   = sum(amt_credit_sum_limit, na.rm = TRUE),
      #max overdue
      bur_max_amt_credit_max_overdue = suppressWarnings(max(amt_credit_max_overdue, na.rm = TRUE)),
      #credit type counts
      bur_n_mortgage         = sum(credit_type == "Mortgage", na.rm = TRUE),
      bur_n_auto             = sum(credit_type == "Car loan", na.rm = TRUE),
      bur_n_consumer_credit  = sum(credit_type == "Consumer credit", na.rm = TRUE),
      bur_n_credit_card      = sum(credit_type == "Credit card", na.rm = TRUE),
      .groups = "drop"
    )
#removes infinity values created from min/max functions and changes to 0, INF and - INF occur if min/max is ran on NA values
bureau_agg <- bureau_agg |> mutate(across(where(is.numeric), ~ ifelse(is.infinite(.x), NA_real_, .x)))
#replace any remaining NA from aggregation with 0
bureau_agg <- bureau_agg |>
  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))

# check for any remaining NAs - none
#bureau_agg |>
  #filter(if_any(everything(), is.na))
```

## Installment Data

```{r installment data}
#take differences prior to aggregation
installment_data <- installment_data |> 
  mutate(day_dif = days_instalment - days_entry_payment # difference between loan is due and when it was paid
        ,amt_dif = amt_instalment - amt_payment # difference between amount of payment due and what was paid
  )

# 2905 rows have NAs, which shows they haven't actually made that payment
colSums(is.na(installment_data[, c("day_dif", "amt_dif")]))
summary(installment_data[, c("day_dif", "amt_dif")])

# view head of data
installment_data |>
  filter(is.na(day_dif) | is.na(amt_dif)) |> head(10)


# remove instalment amounts = to 0
installment_data <- installment_data |>
  filter(amt_instalment != 0)

```

```{r installment data continued}
#below creates calculated fields before aggregation is performed

# set day dif to how long hasn't been paid if no payment
installment_data <- installment_data |> mutate(day_dif = ifelse(is.na(day_dif),days_instalment,day_dif))
# set amt dif to how much is unpaid if no payment
installment_data <- installment_data |> mutate(amt_dif = ifelse(is.na(amt_dif),amt_instalment,amt_dif))
# get ratio of payment difference compared to payment due, flag if paid on time, flag if payment missed
installment_data <- installment_data |> mutate(paid_ratio = amt_dif / amt_instalment,
                                               paid_late = ifelse(day_dif <= 0, 1,0),
                                               missed_payment = ifelse(paid_ratio == 1, 1, 0))
#filter to only have needed columns
installment_data <- installment_data |> select(sk_id_curr, day_dif, amt_dif, paid_ratio, paid_late, missed_payment)
# positive day dif is early payment, 0 is ontime, negative is late
```

```{r installment data aggregation}
#aggregate installment data
installment_agg <- installment_data |>
  group_by(sk_id_curr) |>
  summarise(
    in_n_installments = n(),
    in_avg_day_dif = mean(day_dif, na.rm = TRUE),
    in_min_day_dif = min(day_dif, na.rm = TRUE),
    in_max_day_dif = max(day_dif, na.rm = TRUE),
    in_avg_amt_dif = mean(amt_dif, na.rm = TRUE),
    in_total_amt_dif = sum(amt_dif, na.rm = TRUE),
    in_avg_paid_ratio = mean(paid_ratio, na.rm = TRUE),
    in_share_paid_late = mean(paid_late, na.rm = TRUE),
    in_share_missed_pmt = mean(missed_payment, na.rm = TRUE),
    in_n_paid_late = sum(paid_late, na.rm = TRUE),
    in_n_missed_pmt = sum(missed_payment, na.rm = TRUE),
    .groups = "drop"
  )
#removes infinity values
installment_agg <- installment_agg |> mutate(across(where(is.numeric), ~ ifelse(is.infinite(.x), NA_real_, .x)))
#replace any remaining NA from aggregation with 0
installment_agg <- installment_agg |>
  mutate(across(where(is.numeric), ~ replace_na(.x, 0)))

```

## Credit Card Data

```{r credit card data}
#cleanup NAs
credit_card_data <- credit_card_data |> 
  mutate(AMT_PAYMENT_CURRENT = ifelse(is.na(AMT_PAYMENT_CURRENT),0,AMT_PAYMENT_CURRENT),
         AMT_INST_MIN_REGULARITY =ifelse(is.na(AMT_INST_MIN_REGULARITY),0,AMT_INST_MIN_REGULARITY))

# safe_division funciton to avoid Inf/Na when denominator is 0 or NA
safe_div <- function(num, den) ifelse(is.na(den) | den <= 0, 0, num / den)

#create additional columns
credit_card_data <- credit_card_data |>
  mutate(
    util = safe_div(AMT_BALANCE, AMT_CREDIT_LIMIT_ACTUAL),
    pay_vs_min = safe_div(AMT_PAYMENT_CURRENT, AMT_INST_MIN_REGULARITY),
    pay_to_balance = safe_div(AMT_PAYMENT_TOTAL_CURRENT, AMT_BALANCE), 
    is_delinquent = SK_DPD > 0,                                           
    paid_min = AMT_INST_MIN_REGULARITY > 0 & AMT_PAYMENT_CURRENT >= AMT_INST_MIN_REGULARITY,
    paid_full = AMT_BALANCE > 0 & AMT_PAYMENT_TOTAL_CURRENT >= AMT_BALANCE,      
  )
# check for any NAs
#credit_card_data |>
  #filter(if_any(everything(), is.na))
```

```{r credit card aggregation}
# aggregated data set
credit_card_agg <- credit_card_data |>
  group_by(SK_ID_CURR) |>
  summarise(
    # time covered
    cc_n_months            = n(),
    cc_last_month          = max(MONTHS_BALANCE, na.rm = TRUE),
    cc_span_months         = max(MONTHS_BALANCE, na.rm = TRUE) - 
                          min(MONTHS_BALANCE, na.rm = TRUE),
    # balances, limits, utilization
    cc_avg_balance         = mean(AMT_BALANCE, na.rm = TRUE),
    cc_max_balance         = max(AMT_BALANCE, na.rm = TRUE),
    cc_avg_limit           = mean(AMT_CREDIT_LIMIT_ACTUAL, na.rm = TRUE),
    cc_util_mean           = mean(util, na.rm = TRUE),
    cc_util_max            = max(util, na.rm = TRUE),
    # payments vs due
    cc_pay_vs_min_mean     = mean(pay_vs_min, na.rm = TRUE),
    cc_pay_to_balance_mean = mean(pay_to_balance, na.rm = TRUE),
    # delinquency/behavior
    cc_any_dpd             = any(is_delinquent, na.rm = TRUE),
    cc_share_dpd           = mean(is_delinquent, na.rm = TRUE),
    cc_max_dpd             = max(SK_DPD, na.rm = TRUE),
    cc_share_paid_min      = mean(paid_min, na.rm = TRUE),
    cc_share_paid_full     = mean(paid_full, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    across(where(is.numeric), ~ ifelse(is.infinite(.x), NA_real_, .x))
  )
#forgot to turn to lowercase initially so tidy up
credit_card_agg <- credit_card_agg |>
  rename_with(tolower)
#check for any NA
#credit_card_agg |>
  #filter(if_any(everything(), is.na))
```
## POS Cash Balance

```{r pos cash data}
#cleanup NAs
pos_cash_data <- pos_cash_data |> mutate(CNT_INSTALMENT = ifelse(is.na(CNT_INSTALMENT),0,CNT_INSTALMENT),
                CNT_INSTALMENT_FUTURE = ifelse(is.na(CNT_INSTALMENT_FUTURE),0,CNT_INSTALMENT_FUTURE)
)

# create additional columns
pos_cash_data <- pos_cash_data |>
  mutate(
    # paid and remaining ratios for installments
    share_paid = safe_div(CNT_INSTALMENT - CNT_INSTALMENT_FUTURE, CNT_INSTALMENT),
    share_remaining = safe_div(CNT_INSTALMENT_FUTURE, CNT_INSTALMENT),
    # deliquencies 
    is_dpd = SK_DPD > 0
  )



```


```{r pos cash data agg}
#create aggregation
pos_cash_agg <- pos_cash_data |>
  group_by(SK_ID_CURR) |>
  summarise(
    # duration of months
    pos_n_months        = n(),
    pos_last_month      = max(MONTHS_BALANCE, na.rm = TRUE),
    pos_span_months     = pos_last_month - min(MONTHS_BALANCE, na.rm = TRUE),
    # term and future ratios
    pos_term_median     = median(CNT_INSTALMENT, na.rm = TRUE),
    pos_term_max        = max(CNT_INSTALMENT, na.rm = TRUE),
    pos_future_median   = median(CNT_INSTALMENT_FUTURE, na.rm = TRUE),
    pos_future_min      = min(CNT_INSTALMENT_FUTURE, na.rm = TRUE),
    # repayment progress ratios
    pos_share_paid_mean = mean(share_paid, na.rm = TRUE),
    pos_share_rem_mean  = mean(share_remaining, na.rm = TRUE),
    # delinquency
    pos_any_dpd         = any(is_dpd, na.rm = TRUE),
    pos_share_dpd       = mean(is_dpd, na.rm = TRUE),
    pos_max_dpd         = max(SK_DPD, na.rm = TRUE),
    .groups = "drop"
  ) |>
  # replace inf
  mutate(across(where(is.numeric), ~ ifelse(is.infinite(.x), NA_real_, .x)))

#check for any NA
#pos_cash_agg |>
  #filter(if_any(everything(), is.na))
#forgot to turn to lowercase initially so tidy up
pos_cash_agg <- pos_cash_agg |>
  rename_with(tolower)

```




# Application Data Preperation


## Feature Selection Missing Data and NAs
Variables where greater than 50% of values were missing were dropped. The one exception was ext_source_1. A good portion of those dropped were related to those living in apartments and the characteristics of their housing. 

83 feature variables are retained. 
```{r drop missing data application}
# % missing per variable (full dataset)
missing_perc <- colMeans(is.na(hc_train))

# Keep rule: ≤ 50% missing
cols_to_keep <- names(missing_perc[missing_perc <= 0.50])

# Exception: always keep EXT_SOURCE_1
cols_final <- unique(c(cols_to_keep, "ext_source_1"))

# Filter dataset
hc_train <- hc_train |>
  dplyr::select(all_of(cols_final))

```

## Outliers
One outlier value has been identified and that is the highest income earner listed, this value is removed. 
Also the days employed uses an extreme value for place holding when a person is not employed, so this value is changed from 365,243 to 0. 

```{r remove outliers}
#filter to remove incomes more than 100million
hc_train |> filter(amt_income_total <= 100000000) -> hc_train
#mutate days employed to 0 for positive values
hc_train |> mutate(days_employed = ifelse(days_employed > 0,0,days_employed)) -> hc_train
```



## Joining Data
All four aggregated sets of data are left joined onto application training data set.
```{r merge data}
#join data
hc_merged <- hc_train |> left_join(bureau_agg, by = "sk_id_curr")
hc_merged <- hc_merged |> left_join(installment_agg, by = "sk_id_curr")
hc_merged <- hc_merged |> left_join(credit_card_agg, by = "sk_id_curr")
hc_merged <- hc_merged |> left_join(pos_cash_agg, by = "sk_id_curr")

```


# Training and Test Partions
Splitting into a 80/20 train/test partition split. 

```{r Q3}
set.seed(61)
#create data partition index
index <- createDataPartition(y = hc_merged$target, p = 0.8, list = FALSE)
#create split data frames
train_hc_merged <- hc_merged[index,]
holdout_hc_merged <- hc_merged[-index,] # do not use at all - only for very last run once model picked

```


## Missing Data Imputation and Log Scaling
Imputation of missing values and NAs is done last after data is all joined and split to avoid leakage. Income is log transformed due to it's heavy right skew. Modal values are imputed for categorical, median for numeric. Finally numeric data is scaled to account for certain models being sensitive to magnitude differences of variables. 

A data cleaning recipe is created that can be used on all test and training data. 
For training data an additional recipe is used that creates oversampled data set for use.
```{r recipe creation}
#recipe for data cleaning - to be used on all future data sets
rec <- recipe(target ~ ., data = train_hc_merged) |>
  update_role(sk_id_curr, new_role = "id") |> # set so ID doesn't get touched
  step_zv(all_predictors()) |> #remove zero variance predictors
  step_indicate_na(all_predictors()) |> #flag variable value that was NA, possible signal
  step_impute_median(all_numeric_predictors()) |> #impute medians on NA
  step_impute_mode(all_nominal_predictors()) |> #impute modal for non numeric
  step_mutate(amt_income_total = log1p(amt_income_total)) |> # log scale income - 1p for 0 vals
  step_normalize(all_numeric_predictors(), -starts_with("na_ind_")) |> #scales all numeric - svm/logistic sensitive - doesn't scale flagged NA created values
  step_dummy(all_nominal_predictors(), one_hot = TRUE) |> #one hot encodes all categorical
  step_zv(all_predictors()) #remove zero variance predictors after dummys created

```

Oversampling techniques evaluated
SMOTE - synthetic minority oversampling - creates synthetic nearest neighbor points (two big weaknesses of high n dimensions and categorical data) - for this reason not used

Upsample - randomly oversampling with replacement, most simple and straight forward

ROSE - random oversampling examples - creates synthetic data points using kernel density estimation (small item is it can create nonsensical examples, no other cons except computationally intensive and sensitive to tuning) - for this reason not used but could be revisited


Oversample ratio is used as starting point of .5 (2 to 1 for majority vs minority). This can be tweaked as needed later
```{r over sample recipe}
# baseline recipe for no over sampling
rec_base  <- rec
#recipe for oversampling - adds oversampling at the end of data prep
rec_over_sample <- rec |> step_upsample(target, over_ratio = 0.5, skip = FALSE)
rec_down_sample <- rec |> step_downsample(target, under_ratio = 1, skip = FALSE)
```



```{r apply recipe to data}
#train recipe on training data split
prep_base <- prep(rec_base, training = train_hc_merged, verbose = FALSE)
prep_over <- prep(rec_over_sample, training = train_hc_merged, verbose = FALSE)
prep_down <- prep(rec_down_sample, training = train_hc_merged, verbose = FALSE)
#create data frames
x_train_base <- juice(prep_base)
x_train_over_sample <- juice(prep_over)
x_train_down_sample <- juice(prep_down)

```


# Ensemble Model
An ensemble model is trained on the down sampled data , it uses 5 fold cross validation to determine the optimal hyper parameters. The model cross validation code is lengthy so only ran once and then saved and called in the future. 
```{r train ensemble}
# #rename target for model
# x_train_down_sample2 <- x_train_down_sample |> mutate(target = factor(target, levels = c(0,1), 
#                                                         labels = c("Class0", "Class1"))) |> tidyr::drop_na()
# x_train_base2 <- x_train_base |> mutate(target = factor(target, levels = c(0,1), 
#                                                         labels = c("Class0", "Class1")))
# 
# 
# #create ensemble control
# ensemble_ctrl <- trainControl(
#   method = "cv", number = 5,
#   classProbs = TRUE,
#   summaryFunction = twoClassSummary,
#   savePredictions = "final",
#   allowParallel = TRUE)
# 
# #train model
# ensemble_list <- caretList(
#   target ~ ., data = x_train_down_sample2,
#   trControl = ensemble_ctrl,
#   metric = "ROC",
#   tuneList = list(
#     #will adjust these based on prior optimal xgboost hyper params
#     xgbTree = caretModelSpec(
#       method = "xgbTree",
#       tuneGrid = expand.grid(
#         nrounds = c(5, 10),
#         max_depth = c(3, 5),
#         eta = c(0.05, 0.1),
#         gamma = 0,
#         colsample_bytree = 0.8,
#         min_child_weight = 1,
#         subsample = 0.8)),
#     #will adjust these based on prior optimal nn hyper params
#     nnet = caretModelSpec(
#       method = "nnet",
#       trace = FALSE,
#       preProcess = c("center", "scale"),
#       tuneGrid = expand.grid(
#         size = c(10),
#         decay = c(1e-4, 1e-3)),
#         MaxNWts = 100000,
#         maxit  = 200)))
# 

```


```{r regression ensemble}
#train regression on model stack
# stack_glm <- caretStack(
#   ensemble_list,
#   method = "glm",
#   family = binomial(),
#   metric = "ROC",
#   trControl = ensemble_ctrl)
## model only ran one time and saved below

# cv performance
#trying on downsampled
resamples(ensemble_list) |> summary()

# save ensemble models
#saveRDS(stack_glm, file = "stack_glm_model.rds")
#saveRDS(ensemble_list, file = "ensemble_model.rds")


# load ensemble models
stack_glm <- readRDS("stack_glm_model.rds")
#ensemble_model <- readRDS("ensemble_model.rds")

```

The ensemble model gives an expected AUC ROC of .7417 on the cross validation data.
```{r}

#out of fold AUC - 75%
pred_df <- stack_glm$ens_model$pred   # has obs (factor), and class-prob columns
pos <- levels(pred_df$obs)[1]
stopifnot(pos %in% names(pred_df))
roc_cv  <- pROC::roc(pred_df$obs, pred_df[[pos]], quiet = TRUE)
pROC::auc(roc_cv)
```


# Submission
A final submission for Kaggle is created. This loads in the test data, creates the predictions and generates a file. 
```{r}
#load file - assumes you have copy of file in working directory
hc_test <-read.csv("application_test.csv", stringsAsFactors = TRUE)

#standardize variable names
hc_test <- hc_test |>
  rename_with(tolower)

#convert target and select numeric indicators to factors
hc_test <- hc_test |>
  mutate(
    across(starts_with("flag"),as.factor),
    across(starts_with("reg_"),as.factor),
    across(starts_with("live_"),as.factor)
    )
```



```{r}
#join data
final_merged <- hc_test |> left_join(bureau_agg, by = "sk_id_curr")
final_merged <- final_merged |> left_join(installment_agg, by = "sk_id_curr")
final_merged <- final_merged |> left_join(credit_card_agg, by = "sk_id_curr")
final_merged <- final_merged |> left_join(pos_cash_agg, by = "sk_id_curr")

final_base <- bake(prep_base, new_data = final_merged)
```

```{r}
id_final <- final_base$sk_id_curr
final_base <- final_base |> select(-sk_id_curr) |> as.matrix()
final_pred <- predict(stack_glm, final_base, type = "response")
```

```{r}
output_df <- data.frame(
  SK_ID_CURR = id_final,
  TARGET = as.vector(final_pred)
)

write.csv(output_df, "submission.csv", row.names = FALSE)
```

